---
title: "Course Project: Shiny Application and Reproducible Pitch"
subtitle: "Developing Data Products: Johns Hopkins University Data Science | Coursera"
author: "Hayelom Tesfay"
date: "`r Sys.Date()`"
output:
  ioslides_presentation:
    self_contained: true
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## 1. The Paradigm Shift: Interactive vs. Static

⚠️ The Problem: Traditional Static Reports

- **Fixed Perspectives:** Static tables bury correlations like Weight vs. MPG.
- **Rigid Data:** Manual updates make insights feel stale and disconnected.
- **Blind Spots:** Users cannot drill down into outliers or filter for specifics.

✅ The Solution: A Complete Data Product

- **Total Replacement:** We replace flat images with a live, interactive ecosystem.
- **User-Driven Discovery:** Features like bslib and plotly put the analyst in control.
- **Instant Intelligence:** Dynamic value boxes and automated summaries update in real-time.

🚀 Launch Live App on [**Shinyapps.io**](https://hayelom.shinyapps.io/shiny_pitch_project_submitted/) | 📂 View [**GitHub**](https://github.com/Hayethdt/Developing-Data-Products-Project.git) Source

## 2. Real-Time Predictive Modeling
To power the dashboard, our product calculates Linear Regression on the fly. This allows users to quantify the exact impact of vehicle specs on fuel economy.
```{r regression_summary, echo = TRUE}
# The engine behind the dashboard's trend lines
fit <- lm(mpg ~ hp, data = mtcars)
round(coef(summary(fit)), 4) 
```

✅ Key Insight for the Pitch

The negative coefficient (-0.0682) for hp confirms that for every 10 horsepower increase, efficiency drops by nearly 0.7 MPG. Our app makes this mathematical reality visible instantly.

## 3. Dynamic Data Health & Automated Summaries
Our product doesn't just show data; it audits it. The application processes summary statistics live, ensuring the user always sees the current "Data Health."

```{r summary_code, echo = TRUE}
summary(mtcars$mpg)  # Backend logic executed during user interaction
```
✅ Automated Breakdown:

- **Performance Spread:** A significant 23.5 MPG range across the fleet.
- **Trend Skewness:** Identifying high-efficiency outliers.
- **Instant Context:** Users identify these patterns in seconds.

## 4. Visualizing Trends: From Static to HD

The app leverages ggplot2 and plotly to transform raw numbers into High-Definition insights. Unlike a static image, the live app allows for zooming and hovering to identify specific models.

```{r plot_code, echo = FALSE, fig.height = 2, fig.width = 6}
library(ggplot2)
ggplot(mtcars, aes(x = wt, y = mpg)) + geom_point(color = "#0043CE", size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", formula = y ~ x, color = "#E35205", fill = "#eeeeee") +
  theme_minimal(base_size = 14) + labs(title = "Example Trend: MPG vs. Weight", x = "Weight (1000 lbs)", y = "MPG")
```

### **Conclusion:** Why this Data Product wins:

- *Interactive Tooltips:* Replace labels with deep-dive metadata.
- *Customizable Logic:* Users choose the X-axis; the app handles the math.
- *Professional UI:* Built with bslib for a modern, corporate-ready aesthetic.
